\begin{abstract}
We present a lightweight video face recognition system designed for edge deployment. The core encoder is a MobileNetV4-Conv-Medium student (9.58M parameters, 228M MACs) distilled from a frozen MagFace iResNet-100 teacher (65.2M parameters, 12.15G MACs), achieving a 6.8$\times$ parameter reduction and 53.3$\times$ MACs reduction. The student is trained with a combination of MagFace classification loss and embedding-level knowledge distillation. We study three independently trained variants: a clean KD baseline (V1), a spatial KD variant with aggressive occlusion curriculum (V2), and an SWA-stabilized robust variant (V3). The clean KD student achieves 87.98\% TAR at FAR=$10^{-4}$ on IJB-B and 90.65\% on IJB-C, within 1.44\% of MobileFaceNet (W600K) on IJB-B and within 1.07\% on IJB-C, despite using 7$\times$ fewer training identities. The SWA variant substantially reduces degradation under synthetic lower-face masking (TAR drop reduced by up to 52\% relative to V1 on CALFW-Occ) while maintaining strong verification performance. On the RMFRD real masked-face dataset (RetinaFace-aligned), the SWA variant (84.18\% AUC, 23.03\% Rank-1) outperforms the clean KD variant (82.66\%, 22.01\%), confirming that occlusion robustness transfers to real-world masks. On the MFR2 dataset (53 identities, mixed masked/unmasked pairs), both student variants substantially outperform MobileFaceNet on verification AUC (V3/SWA: 94.38\% vs.\ MBF: 89.91\%), and V3/SWA leads on 1:N identification AUC (95.99\% vs.\ 95.11\%), though MBF retains a Rank-1 advantage attributed to its larger training set. Crucially, V3/SWA outperforms MobileFaceNet (W600K) on all five synthetic occlusion benchmarks despite MBF being trained on $7\times$ more identities: MBF drops up to $-0.550$ TAR@$10^{-3}$ on AgeDB under masking, while V3/SWA drops only $-0.321$ — a $41.6\%$ reduction. The distilled encoder is integrated into a complete runtime pipeline with YOLO-based face detection, 5-point alignment, multi-object tracking, liveness filtering, MagFace magnitude-based quality gating, magnitude-weighted track-level pooling, and FAISS-based identity retrieval. A template pooling ablation on IJB-B/C confirms that magnitude-weighted pooling consistently outperforms simple mean pooling at strict FAR operating points. Limitations include the absence of standardized open-set 1:N video evaluation and a fully controlled PAD ablation, which are left for future work.
\end{abstract}
