\section{Results}

\subsection{Clean Image Verification}

Table~\ref{tab:bin_results} reports the verification accuracy on LFW, CFP-FP, and AgeDB-30\cite{cplfw,cfp,agedb}. The teacher model achieves the highest overall accuracy, as expected from the larger MagFace iResNet-100 backbone\cite{magface}. Among student variants, V1/best achieves the highest mean accuracy (96.37\%), driven by strong CFP-FP and AgeDB-30 scores. V4/best — the spatial-KD variant that applies spatial feature matching during clean epochs and masks during curriculum epochs — achieves the highest LFW accuracy among all students (99.58\%), but with lower CFP-FP and AgeDB-30 scores than V1.

\begin{table}[h]
\centering
\caption{Clean face verification results on LFW, CFP-FP, and AgeDB-30.}
\label{tab:bin_results}

\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccc}
\hline
\textbf{Model} & \textbf{LFW} & \textbf{CFP-FP} & \textbf{AgeDB-30} & \textbf{Mean} \\
\hline
MagFace iResNet-100 & 99.78 & 96.27 & 98.30 & 98.12 \\
V1/latest & 99.25 & 93.43 & 95.68 & 96.12 \\
V1/best & 99.20 & \textbf{94.14} & \textbf{95.77} & \textbf{96.37} \\
V3/best & \textbf{99.43} & 92.79 & 95.15 & 95.79 \\
V3/latest & 99.02 & 92.06 & 93.60 & 94.89 \\
V3/SWA & 99.00 & 91.54 & 94.00 & 94.85 \\
V4/best & \textbf{99.58} & 92.71 & 95.32 & 95.87 \\
\hline
\end{tabular}
}
\end{table}

These results confirm a clear accuracy-robustness trade-off. V1 is the strongest student for clean-face verification by mean accuracy. V4/best achieves the best LFW score (99.58\%) across all students, showing that spatial KD during clean training epochs improves frontal-face discriminability, but at a cost to CFP-FP (profile-pose) performance. V3 variants trade further clean accuracy for occlusion robustness.

\subsection{IJB Template Verification}

Table~\ref{tab:ijb_results} reports template-based verification results on IJB-B and IJB-C. We focus on TAR@$10^{-4}$ because this operating point is more relevant for security-sensitive deployment than average accuracy alone\cite{ijbb,ijbc}.

\begin{table*}[h]
\centering
\caption{IJB-B and IJB-C template verification results. TAR is reported at FAR=$10^{-4}$.}
\label{tab:ijb_results}
\begin{tabular}{lcccc}
\hline
\textbf{Model} & \textbf{IJB-B AUC} & \textbf{IJB-B TAR@$10^{-4}$} & \textbf{IJB-C AUC} & \textbf{IJB-C TAR@$10^{-4}$} \\
\hline
Teacher: MagFace iResNet-100 & 0.9922 & 93.14 & 0.9960 & 97.64 \\
V1/latest & 0.9912 & \textbf{87.98} & 0.9937 & \textbf{90.65} \\
V1/best & \textbf{0.9938} & 86.88 & 0.9929 & 89.50 \\
V2/latest & 0.9935 & 84.52 & 0.9949 & 86.85 \\
V3/best & 0.9916 & 84.77 & 0.9929 & 87.64 \\
V3/latest & 0.9917 & 84.78 & 0.9932 & 87.23 \\
V3/SWA & 0.9919 & 85.27 & 0.9930 & 87.77 \\
V4/best & 0.9933 & 84.35 & \textbf{0.9953} & 87.35 \\
\hline
\end{tabular}
\end{table*}

The teacher remains the strongest model on both IJB-B and IJB-C. Among the student models, V1/latest gives the best low-FAR performance (87.98\% TAR@$10^{-4}$ on IJB-B; 90.65\% on IJB-C), confirming V1 as the preferred checkpoint for clean or controlled deployment. V1/best achieves the highest student AUC on IJB-B (0.9938), slightly above V4/best (0.9933) and V2/latest (0.9935). V4/best reaches 84.35\% TAR@$10^{-4}$ on IJB-B — between V3/SWA (85.27\%) and V3/best (84.77\%) — despite having the best student LFW accuracy. On IJB-C, V4/best achieves the highest student AUC (0.9953) and 87.35\% TAR@$10^{-4}$, surpassing V2/latest (0.9949) and all V3 variants on AUC. The gap between V4/best's LFW advantage and its IJB-B TAR@$10^{-4}$ result reflects that spatial KD improves frontal-face discriminability but does not generalize as strongly to pose-diverse IJB-B probes; the stronger V4/best IJB-C AUC suggests better overall template calibration at moderate FAR thresholds.

V2/latest obtains a high AUC (0.9935) but weaker TAR@$10^{-4}$ (84.52\%), illustrating that AUC alone does not fully reflect strict low-FAR behavior. Real access-control deployments operate at strict FAR thresholds where TAR metrics are more relevant.

\subsection{Occlusion Robustness}

Table~\ref{tab:occlusion_results} reports the TAR@$10^{-3}$ drop under synthetic lower-face masking. A smaller absolute drop indicates better robustness. We include MobileFaceNet (W600K) to provide a direct comparison against the lightweight baseline.

\begin{table}[h]
\centering
\caption{Occlusion robustness: TAR@$10^{-3}$ drop (masked$-$clean) under synthetic lower-face masking. Smaller absolute value is better.}
\label{tab:occlusion_results}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccccc}
\hline
\textbf{Model} & \textbf{LFW} & \textbf{CFP-FP} & \textbf{AgeDB} & \textbf{CPLFW} & \textbf{CALFW} \\
\hline
MagFace iResNet-100 & -0.041 & -0.520 & -0.541 & -0.724 & -0.236 \\
MobileFaceNet (W600K) & -0.085 & -0.549 & -0.550 & -0.494 & -0.324 \\
V1/latest & -0.037 & -0.369 & -0.462 & -0.222 & -0.248 \\
V4/best & -0.036 & -0.338 & -0.344 & -0.329 & -0.225 \\
V3/SWA & \textbf{-0.019} & \textbf{-0.298} & \textbf{-0.321} & \textbf{-0.148} & \textbf{-0.100} \\
\hline
\end{tabular}
}
\end{table}

All three student variants outperform MobileFaceNet (W600K) on most occlusion benchmarks, despite MBF being trained on $7\times$ more identities. MBF suffers the largest TAR drops: $-0.549$ on CFP-FP, $-0.550$ on AgeDB, and $-0.494$ on CPLFW — expected because MBF has no occlusion augmentation. V4/best (spatial KD, partial masking curriculum at epoch 23) is intermediate: it substantially outperforms V1/latest on LFW, CFP-FP, AgeDB, and CALFW, but shows worse CPLFW drop ($-0.329$ vs $-0.222$) — likely because spatial KD trades some pose-invariance for frontal-face discriminability. V3/SWA remains the most robust across all five protocols.

V3/SWA reduces the absolute TAR drop relative to MBF by $45.7\%$ on CFP-FP, $41.6\%$ on AgeDB, $70.1\%$ on CPLFW, and $69.1\%$ on CALFW. V4/best achieves intermediate improvements of $38.4\%$, $37.3\%$, $33.5\%$, and $30.6\%$ respectively, consistent with partial masking-curriculum completion at the selected checkpoint (epoch~23 of 40, $\hat{p}_{\text{mask}}\approx0.20$).

The teacher also suffers large degradation under lower-face masking, especially on CFP-FP, AgeDB, and CPLFW. This shows that neither model scale nor training data volume guarantees robustness to structured occlusion\cite{cplfw,cfp,agedb,izmailov2019averagingweightsleadswider}.

\subsection{Overall Comparison}

Table~\ref{tab:checkpoint_recommendation} summarizes the practical role of each checkpoint. V1/latest is the best choice for clean deployment because it gives the strongest student performance on IJB-B and IJB-C at strict FAR. V1/best is preferable when the goal is maximum bin-protocol accuracy. V3/SWA is the best choice when lower-face occlusion or mask usage is expected. V2/latest is kept as an ablation rather than a recommended deployment checkpoint.

\begin{table*}[h]
\centering
\caption{Recommended use cases for the evaluated checkpoints.}
\label{tab:checkpoint_recommendation}
\begin{tabular}{lll}
\hline
\textbf{Checkpoint} & \textbf{Strength} & \textbf{Recommended Use} \\
\hline
V1/latest & Best student IJB TAR@$10^{-4}$ & Clean access-control deployment \\
V1/best & Best bin-protocol mean accuracy & Clean image-pair verification \\
V2/latest & Diagnostic spatial-KD ablation & Not recommended for deployment \\
V3/latest & Robust variant before averaging & Reference checkpoint \\
V3/SWA & Best occlusion robustness & Masked or partially occluded deployment \\
V4/best & Best student LFW (99.58\%) and IJB-C AUC (0.9953); intermediate occlusion robustness & Frontal-face-heavy controlled environments \\
\hline
\end{tabular}
\end{table*}

Overall, the results reveal a spectrum of trade-offs rather than a single best model. V1/latest is the strongest student for clean IJB template verification at strict FAR; V1/best leads on frontal-face bin protocols by mean accuracy and IJB-B AUC; V4/best leads on LFW (99.58\%), IJB-C AUC (0.9953), and MFR2 Rank-1 (69.01\%); and V3/SWA provides the best resistance to lower-face occlusion. V4/best occupies an intermediate position — better than V1 on occlusion robustness (partial masking curriculum), and better than V3 on LFW and IJB-C AUC (spatial KD during clean epochs). This hierarchy confirms that the appropriate deployment checkpoint depends on whether the operational environment is predominantly clean-face, pose-diverse, or mask-occluded.

\subsection{Computational Efficiency}

Table~\ref{tab:efficiency} summarizes the computational characteristics of the teacher, student, and MobileFaceNet baseline. All measurements were taken on a single Tesla P100 16~GB at batch~1 for latency and batch~64 for throughput.

\begin{table}[h]
\centering
\caption{Computational efficiency comparison. Latency and throughput measured on Tesla P100.}
\label{tab:efficiency}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccc}
\hline
\textbf{Model} & \textbf{Params} & \textbf{MACs} & \textbf{Latency (ms)} & \textbf{Memory} \\
\hline
MagFace iResNet-100 & 65.2M & 12.15G & 18.52 & 249 MB \\
\textbf{MobileNetV4-M (student)} & \textbf{9.58M} & \textbf{228M} & \textbf{11.22} & \textbf{38.3 MB} \\
MobileFaceNet (W600K) & $\sim$1M & $\sim$224M & — & — \\
\hline
\end{tabular}
}
\end{table}

The student achieves a 6.8$\times$ parameter reduction and 53.3$\times$ MACs reduction relative to the teacher, while also being faster in wall-clock latency (11.22~ms vs.~18.52~ms at batch~1). The memory footprint is reduced from 249~MB to 38.3~MB, making the student suitable for edge deployment on resource-constrained devices.

\subsection{Lightweight Baseline Comparison}

To contextualize the student's efficiency, we compare against MobileFaceNet trained on WebFace600K (\textasciitilde600K identities) using the insightface \texttt{buffalo\_sc} model with ArcFace loss. Our student uses MS1M-RetinaFace (\textasciitilde85K identities) — 7$\times$ fewer training identities.

\begin{table}[h]
\centering
\caption{Lightweight model comparison on IJB-B/C template verification at FAR=$10^{-4}$.}
\label{tab:baseline_comparison}
\begin{tabular}{lccc}
\hline
\textbf{Model} & \textbf{Params} & \textbf{IJB-B TAR@$10^{-4}$} & \textbf{IJB-C TAR@$10^{-4}$} \\
\hline
Teacher: iResNet-100 & 65.2M & 93.14 & 97.64 \\
MobileFaceNet (W600K) & $\sim$1M & 89.42 & 91.72 \\
Student: V1/latest & 9.58M & 87.98 & 90.65 \\
Student: V4/best & 9.58M & 84.35 & 87.35 \\
Student: V3/SWA & 9.58M & 85.27 & 87.77 \\
\hline
\end{tabular}
\end{table}

MobileFaceNet (W600K) has $\sim$1M parameters and $7\times$ more training identities, yet our 9.58M-parameter V1/latest student is within 1.44\% on IJB-B and within 1.07\% on IJB-C. V4/best (84.35\% IJB-B TAR@$10^{-4}$; 87.35\% IJB-C TAR@$10^{-4}$) trails V1/latest by 3.63\% on IJB-B but achieves the highest student IJB-C AUC (0.9953), consistent with spatial KD improving overall template calibration while not fully recovering the IJB-B low-FAR advantage. The key advantage of all student variants over MBF lies in occlusion robustness, where MBF is significantly weaker (see RMFRD and occlusion evaluations).

\subsection{Real-World Masked Face Evaluation (RMFRD)}

To evaluate robustness on real masked faces beyond synthetic lower-face masking, we use the RMFRD/AFDB dataset: 403 paired identities with clean gallery images and real-world masked probe images. Images are pre-processed using InsightFace's RetinaFace (\texttt{det\_500m}) detector to obtain 5-point face landmarks and apply standard ArcFace affine alignment to 112$\times$112.

\textbf{Protocol:} Gallery = mean embedding of up to 10 clean images per identity; Probes = all masked images; Positive pairs = same-identity mask-to-clean matching; TAR@FAR and Rank-1 identification are reported.

\begin{table}[h]
\centering
\caption{Real-world masked face recognition on RMFRD (403 paired identities, RetinaFace-aligned). Gallery: mean of up to 10 clean images per identity. Probe: 1,945 real masked images. Fallback bilinear resize for 7.9\% of masked images where RetinaFace detected no face. Unaligned reference: MBF 68.72\% AUC / 7.25\% Rank-1; V1/best 71.02\% / 4.78\%; V3/SWA 69.90\% / 6.84\%.}
\label{tab:rmfrd}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccc}
\hline
\textbf{Model} & \textbf{AUC} & \textbf{TAR@$10^{-3}$} & \textbf{Rank-1} \\
\hline
MobileFaceNet (W600K) & \textbf{85.55} & \textbf{26.43} & \textbf{36.56} \\
Student: V1/best & 82.66 & 15.06 & 22.01 \\
Student: V4/best & 82.07 & 8.02 & 20.67 \\
Student: V3/SWA & 84.18 & 15.68 & 23.03 \\
\hline
\end{tabular}
}
\end{table}

The alignment step is critical for meaningful evaluation. Without alignment, all models use simple bilinear resize, which suppresses discriminative facial features: MBF achieves only 68.72\% AUC / 7.25\% Rank-1 on unaligned crops but reaches 85.55\% AUC / 36.56\% Rank-1 after RetinaFace alignment — an improvement of $+16.8\%$ AUC and $+29.3$ Rank-1 points. This confirms that the RMFRD raw images are not pre-aligned and that any meaningful comparison must use landmark-based preprocessing.

Among student models, V3/SWA (84.18\% AUC, 23.03\% Rank-1) outperforms both V1/best (82.66\%, 22.01\%) and V4/best (82.07\%, 20.67\%) on real masked probes. The ordering V3/SWA $>$ V1/best $\approx$ V4/best on RMFRD AUC is consistent with the synthetic masking ablation (Table~\ref{tab:occlusion_results}): the full occlusion curriculum and SWA stabilization confer real-world masked face robustness, while V4/best at epoch~23 (partial curriculum) matches but does not exceed V1/best. The low V4/best TAR@$10^{-3}$ (8.02\%) reflects the sensitivity of strict FAR metrics to the hard-negative distribution in a small dataset (9,725 negative pairs).

MobileFaceNet (W600K), despite its smaller parameter count (1M vs.\ 9.58M), outperforms both student variants on this dataset (85.55\% AUC, 36.56\% Rank-1). We attribute this primarily to the significantly larger training set: MBF is trained on WebFace600K with $\sim$600K identities, while our student uses MS1M-RetinaFace with $\sim$85K identities (7$\times$ fewer). Additionally, the RMFRD dataset predominantly features Asian celebrities, who may be more represented in WebFace600K than in MS1M. This suggests that the student's IJB advantage does not fully translate to this particular domain, highlighting the importance of training data diversity for cross-domain robustness.

\subsection{Real-World Masked Face Evaluation (MFR2)}

MFR2 is a real-world masked face dataset of 55 celebrity identities with both masked and unmasked images per subject. We evaluate 53 identities that have at least one clean (\texttt{no-mask}) and one masked image. Two protocols are evaluated:

\begin{itemize}
    \item \textbf{Verification (pairs.txt)}: 424 same-identity and 424 different-identity pairs sourced from the official LFW-style split; pairs may combine masked and unmasked images. We report AUC.
    \item \textbf{1:N Identification}: gallery = mean embedding of all clean images per identity (98 total); probes = all masked images (171 total). We report AUC and Rank-1.
\end{itemize}

Note: With only 424 negative pairs in the verification split, the FAR resolution ($\approx 1/424 \approx 0.0024$) is too coarse to report meaningful TAR at FAR$\leq10^{-3}$; we therefore report AUC only for MFR2.

\begin{table}[h]
\centering
\caption{MFR2 evaluation (53 paired identities). Verification: pairs.txt AUC (mixed masked/clean pairs). Identification: 1:N AUC and Rank-1 with clean gallery and masked probes.}
\label{tab:mfr2}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccc}
\hline
\textbf{Model} & \textbf{Verif AUC} & \textbf{ID AUC} & \textbf{ID Rank-1} \\
\hline
MobileFaceNet (W600K) & 89.91 & 95.11 & \textbf{72.51} \\
Student: V1/best      & 94.18 & 95.37 & 61.99 \\
Student: V4/best      & 94.28 & 95.48 & 69.01 \\
Student: V3/SWA       & \textbf{94.38} & \textbf{95.99} & 66.67 \\
\hline
\end{tabular}
}
\end{table}

All three student variants substantially outperform MobileFaceNet (W600K) on MFR2 verification AUC (V1/best: $+4.27\%$; V4/best: $+4.37\%$; V3/SWA: $+4.47\%$). V4/best ranks second on identification AUC (95.48\%) and achieves the highest student Rank-1 (69.01\%), surpassing V3/SWA (66.67\%) and V1/best (61.99\%). This is consistent with V4's stronger frontal-face discriminability (LFW 99.58\%) translating to gallery-matching accuracy on clean celebrity images. V3/SWA leads on identification AUC ($+0.88\%$ over MBF), confirming broader embedding calibration.

MBF retains the highest Rank-1 identification ($72.51\%$), attributed to its larger training set ($\sim\!600$K identities vs.\ $\sim\!85$K) and likely greater overlap with the post-2020 celebrity identities in MFR2. The student variants' superior AUC suggests better relative pairwise discriminability despite fewer absolute top-1 successes.

\subsection{Runtime Pipeline Results}

The runtime pipeline was designed to evaluate the trained encoder under realistic video conditions. It records processed frames, detection counts, accepted observations, recognized observations, match rejection reasons, and mean FPS. However, since no standardized annotated video benchmark is used in the current evaluation, we do not report a single comparable runtime accuracy table in this section. Instead, the benchmark results above focus on the recognition quality of the trained encoders, while the runtime pipeline is used to validate practical deployment behavior with detection, tracking, liveness filtering, quality gating, pooling, and FAISS retrieval\cite{faiss}.