\section{Conclusion}

This work presents an edge-oriented video face recognition system based on a lightweight MobileNetV4 student distilled from a frozen MagFace iResNet-100 teacher. The proposed system combines MagFace classification, embedding-level knowledge distillation, and a complete runtime pipeline with detection, tracking, liveness filtering, magnitude-based quality gating, track-level pooling, and FAISS-based retrieval\cite{mobilenetv4}\cite{relationalkd}\cite{hinton2015distilling}\cite{faiss}\cite{botsort}\cite{deepsort}.

We evaluate four independently trained variants rather than sequential training phases. V1 is the best choice for clean-face verification under strict low-FAR IJB evaluation, and V3/SWA provides the best robustness under lower-face occlusion. V4/best — applying spatial KD only during clean epochs with a masking curriculum — achieves the highest student LFW accuracy (99.58\%), the highest student IJB-C AUC (0.9953), the highest student MFR2 Rank-1 (69.01\%), and intermediate occlusion robustness, confirming that gated spatial KD improves frontal-face discriminability and overall template calibration while partial masking training provides some occlusion generalization. V2 serves as an ablation, showing that un-gated spatial KD during masked epochs can hurt performance\cite{izmailov2019averagingweightsleadswider}\cite{ijbb}\cite{ijbc}\cite{faiss}.

Overall, the study highlights a practical trade-off between clean accuracy and occlusion robustness. For controlled access-control scenarios, the clean KD variant is preferred. For environments with masks or partial occlusion, the SWA-stabilized robust variant is more suitable. Future work will focus on stronger anti-spoofing evaluation, real open-set video benchmarks, and further optimization for real-time edge deployment.

\section{Limitations}

This work has several limitations. First, the runtime video pipeline has not yet been evaluated on a standardized large-scale open-set video benchmark. Although the system records runtime statistics such as FPS, accepted observations, recognized observations, and match rejection reasons, a more complete evaluation with annotated video tracks is needed to measure TPIR, FPIR, FNIR, and identity switches.

Second, the anti-spoofing component is treated as a modular liveness gate rather than a fully optimized presentation attack detection system. Its effectiveness depends on the selected PAD model and the availability of real spoofing datasets. Therefore, stronger evaluation on real presentation attack protocols is required before deployment in high-security authentication scenarios.

Third, the occlusion robustness evaluation is based on synthetic lower-face masking. While this protocol is useful for controlled comparison, it does not fully represent real-world occlusions such as masks with different shapes, hands, scarves, extreme pose, motion blur, or low-resolution surveillance footage.

Fourth, the training variants are practical configuration-level ablations rather than fully isolated single-factor experiments. Since spatial KD gating, occlusion strength, and SWA differ across variants, future work should isolate each factor more carefully. In particular, the V4 SWA phase failed due to MagFace embedding-norm inflation under the heavy masking curriculum (norms $>10^3$ at epochs 35--39 vs.\ $\sim\!13$ for the best checkpoint); the selected V4/best checkpoint at epoch~23 represents only partial masking-curriculum completion ($\hat{p}_{\text{mask}}\approx0.20$). A norm-clamped or norm-regularized retraining may yield a stronger V4 final model.

\section{Broader Impact and Ethical Considerations}

Face recognition technology carries important ethical implications that must be acknowledged. First, regarding privacy: the deployment of automated face recognition in public or semi-public environments raises privacy concerns, especially without the knowledge or consent of the individuals being recognized. We recommend that any deployment of this system respect applicable data protection regulations (e.g., GDPR) and obtain explicit consent from subjects whenever feasible.

Second, regarding demographic fairness: training datasets for face recognition, including MS1M, are known to have demographic imbalances in terms of age, gender, and ethnicity. We did not conduct demographic subgroup analyses in this work. It is plausible that performance varies across demographic groups, and such disparities could amplify existing biases in access-control contexts. Future work should evaluate recognition accuracy disaggregated by demographic subgroup and apply appropriate debiasing strategies.

Third, regarding misuse: the pipeline described in this work is intended for controlled access-control scenarios (e.g., office or campus entry systems). Deployment in mass surveillance, tracking of individuals without consent, or any scenario that infringes on civil liberties is not intended and is strongly discouraged. The open-set unknown-identity grouping feature, in particular, should not be used to build covert profiles of private individuals.